
Financial Evaluation and Investment Strategy   
Portfolio Optimization   
Assignment 1.7


```{r cache=TRUE}


library(quadprog)
library(stockPortfolio)

stocks <- c(
 "SPY",
 "EFA",
 "IWM",
 "VWO",
 "LQD",
 "HYG")

returns <- getReturns(stocks, freq="week")
# returns$R - matrix(*,6) - cols are the stocks, rows are the weekly returns 

# covariance of a stock with itself is just its variance.
returns.cov <- cov(returns$R)

# correlation of a stock with itself is always 1.
returns.cor <- cor(returns$R)

spy.var <- var(returns$R[,"SPY"])
spy.sd <- sqrt(spy.var)
efa.var <- var(returns$R[,"EFA"])
efa.sd <- sqrt(efa.var)

# relationship between covariance and correlation
returns.cov["SPY","SPY"] / spy.sd / spy.sd
returns.cov["SPY","EFA"] / spy.sd / efa.sd
returns.cor["SPY","EFA"] 


```


### Question 1

What's the min port var / port ret / sharpe when all weights = 0.25 ?

```{r}

stocks.covar <- matrix( c(
                            0.0369,0.0522,0.0513,0.0406,
                            0.0522,0.1552,0.1105,0.0576,
                            0.0513,0.1105,0.1116,0.0526,
                            0.0406,0.0576,0.0526,0.0506 ),
                        nrow=4 )
stocks.er <- matrix( c( 
                        11.20,
                        19.00,
                        17.50,
                        11.10 ) / 100,
                     nrow=4)
colnames(stocks.er) <- c("large", "small", "value", "growth")

stocks.sd <- c( 
                19.20,
                39.40,
                33.40,
                22.50) / 100 
colnames(stocks.sd) <- c("large", "small", "value", "growth")

## ------------------------------------------------
## What's the min port var / port ret / sharpe when all weights = 0.25 ?
## ------------------------------------------------

weights <- matrix( c(0.25, 0.25, 0.25, 0.25), nrow=4 )

port.var <- t(weights) %*% stocks.covar %*% weights
port.var
port.sd <- sqrt(port.var)
port.sd

port.ret <- t(weights) %*% stocks.er
port.ret

riskfree.ret <- 3.5 / 100
port.sharpe <- (port.ret - riskfree.ret) / port.sd
port.sharpe
```



### Question 1b

What's the max port return with similar risk as equally-weighted portfolio above, given you CANNOT short?

```{r}

source("eff.frontier.R")

stocks.er.vec <- t(stocks.er)
colnames(stocks.er.vec) <- c("large", "small", "value", "growth")
eff.noshort <- eff.frontier(stocks.covar, stocks.er.vec , short="no", risk.free.rate=3.5/100)
eff.noshort <- round(eff.noshort,6)

# format(eff.noshort, digits=5, scientific=6)
eff.noshort

```

### Question 2

What's the max port return with similar risk as equally-weighted portfolio above, given you CAN short?

```{r}
eff.short <- eff.frontier(stocks.covar, stocks.er.vec, short="yes", risk.free.rate=3.5/100)

eff.short <- round(eff.short, 6)
eff.short

eff.varEquals(eff.short, 0.2603)

```


### Question 3

Find the weights that produce the min port variance, given you CANNOT short.

```{r}

# eff.noshort[ which(eff.noshort$Std.Dev == min(eff.noshort$Std.Dev)), ]
eff.minVariance( eff.noshort )


```

### Question 4

Find the weights that produce the max sharpe ratio, given you CANNOT short.


```{r}

# eff.noshort[ which(eff.noshort$sharpe == max(eff.noshort$sharpe)), ]
eff.maxSharpe( eff.noshort )

```

