
## Fin-Eval: Module 3.2 In-class Assignment

Rob Alderman  
Financial Evaluation and Strategy  
Module 3.2  
Sept 2015  

### Objective

Analyze the performance of value stocks (HML) from 1927 - 2014 in a CAPM framework. 


### Load data

```{r}
Sys.setenv(JAVA_HOME="C:\\fox\\java\\jre")
p <- Sys.getenv("PATH")
p1 <- paste("C:\\fox\\java\\jre\\bin\\j9vm;", p, sep="")
Sys.setenv(PATH=p1)

library(xlsx)

data <- read.xlsx("HML_1927_2014.xlsx", sheetIndex=1, rowIndex=4:1060,colIndex=1:4)
str(data)
head(data)



### Q1.  What's the average monthly return for HML?

```{r}
with(data, mean(HML))

```


### Q2.  How much of HML return reflects performance beyond its benchmark (alpha)?

```{r}

model.hml <- lm(HML ~ Mkt_RF, data=data)
summary(model.hml)
model.hml.alpha <- coef(model.hml)["(Intercept)"]
model.hml.beta <- coef(model.hml)["Mkt_RF"]

library(ggplot2)
qplot(y=HML, x=Mkt_RF, data=data) +
    ggtitle("CAPM - HML") + 
    geom_hline(y=0,colour="lightgray") +
    geom_vline(x=0,colour="lightgray") +
    geom_abline(intercept = model.hml.alpha, 
                slope = model.hml.beta, 
                colour="blue")

# % of HML return that reflects outperformance of benchmark:
model.hml.alpha / mean(data$HML)
```

### Q3.  How much of HML return reflects systemic risk (beta)?

```{r}

(mean(data$HML) - model.hml.alpha) / mean(data$HML)

```


### Q4.  Does HML perform better in good times or bad?  Seems like good.



### Q5. What fraction of HML variability is explained by market-wide movements? (R<sup>2</sup>)

```{r}
summary(model.hml)$r.squared
```



### 6. Have the publications of papers on SMB (in 1980) and HML (in 1992/93) affected the returns for those asset classes?

```{r}

data.1980.down <- subset(data, date < 198001)
data.1980.up <- subset(data, date >= 198001)

data.1992.down <- subset(data, date < 199201)
data.1992.up <- subset(data, date >= 199201)

mean(data$Mkt_RF)
mean(data.1980.down$Mkt_RF)
mean(data.1980.up$Mkt_RF)
mean(data.1992.down$Mkt_RF)
mean(data.1992.up$Mkt_RF)

mean(data$HML)
mean(data.1992.down$HML)
mean(data.1992.up$HML)  # after Fama's paper

with(data, mean(SMB))
with(data.1980.down, mean(SMB))
with(data.1980.up, mean(SMB))
with(data.1992.up, mean(SMB))


model.hml.1992 <- lm(HML ~ Mkt_RF, data=data.1992.up)
summary(model.hml.1992)
```

